### 2016-10-26 doc/publications/plos16/plos16.tex
Fix typos and things

\end{flushleft}
\tableofcontents
\clearpage
\section{Notes}
\begin{itemize}
\item I think the structure of the first section should be
  \begin{itemize}
  \item convergence results from Biau 2012  (done)
  \item consider  \cite{Segal.2004} examples
  \item simulations reproduced from Genuer et al 2010 (no -- we need classification examples)
  \item check Piotr's simulation  (done)
  \item run sparkML or CursedForest on the  simulation data 
  \item we extend the simulations to many more variables 
  \item I dont think I can get R to handle chromosme 1 etc. May have to run CF on chromosme 22, see table
    \ref{table:smaller_data} -- On it, Aidan
  \item variable  selection may succeed where accurate prediction does not
  \item discussion of the effect of the \ntree\ and \mtry\ parameters for variable selection with many noise variables
  \item   perhaps we argue that in the case where the relationship of $X$ and $y$ takes a particular form (step, ramp) we can recover
    the significant variables in the presence of a very large number of noise variables
  \item in the case of genomic data, our functions of interest are more likely to be stop or ramp functions rather than harmonic
    functions. So we have some hope.
  \end{itemize}
\item check out 
  \begin{itemize}
  \item D\'iaz-Uriarte, R., \& Alvarez de Andr\'es, S. (2006). (done)
  \item \cite{Goldstein.et.al.2011}
notes that the 0, 1, 2 encoding of alleles is an additive model i.e. each additional minor allele increases the effect 
\begin{verbatim}
Different genetic effects
Type	Mechanism	Partition
Additive	Each additional minor allele increases variation	0, 1, 2
Dominant	Presence of at least 1 minor allele increases variation	0, 1/2
Recessive	Two minor alleles needed for variation	0/1, 2
Heterosis	Heterozygote leads to variation	0/2, 1
\end{verbatim}

@@

\marginpar{decison trees capture interaction -- there is no main effect term. Random forests inherit this property. I dont think it
  has anything to do with boostrapping, the point of which is to ensure that successive trees are not correlated.}
Random forests~\cite{Breiman2001} are able to capture interactions between features while still allowing to
prioritise SNPs for their contribution to disease.  The joint interaction is achieved by appling the technique of
bootstrap aggregation or bagging to decision trees, while a measure of variable importance can be provided, for example,
by summing the number of times a variable is used to split a node, over all nodes and all trees.  Random forests are well suited for
processing ``wide'' genomic data for two reasons.  Firstly, while other machine learning applications have the
propensity to overfit datasets with more features $p$ than samples $n$ (a consequence of the ``curse of
dimensionality''~\cite{Bauer2014, bellman1961adaptive}), decision trees are resistant to overfitting. 
Secondly, random forests are also very easy to parallelise. As the forest is a sum of decision trees, it it possible to grow
separate trees on different processors and combine the trees.

The most popular random forest software packages are R-based, although current implementations do not scale well with
increasing feature size~\cite{Wright.and.Ziegle.2016}.  This is despite R now supporting long vectors (length greater
than $2^{31}$), and the {\sc randomForest} implementation in R lending itself to paralelization to grow multiple trees
simultaneously~\cite{Liaw.and.Weiner.2002} and them combine them.  More suitable implementations for ``wide'' genomic data have been
developed in other programming languages.  {\sc CloudForest}~\cite{Bressler2015} written in Go, achieves fast running
times by effective use of the CPU cache, optimising for different classes of features and efficiently multi-threading.
Similarly, {\sc Ranger}~\cite{Wright.and.Ziegle.2016} written in C++ with an R front-end is specifically optimised for
dealing with large data sets.

@@

Since the original implementation of random forests, there have been a number of developments leading to a better understanding of
the algorithm. 
\begin{itemize}
\item random forests are resistant to overfitting. However it is not true that they will not overfit,
  ~\cite{Segal.2004} gives examples of overfitting in the case where:
  \begin{itemize}
  \item a single tree is the most appropriate model; and
  \item the variables are hightly correlated so that the boostrap resampling does not give rise to uncorrelated trees.
  \end{itemize}
However, keeping these difficult cases in mind, random forests have a strong claim to being resistant to overfitting.

\item \cite{Strobl.et.al.2007} demonstrates that the random forest algorithm is subject to a bias in variable selection via two mechanisms:
  \begin{enumerate}
  \item differences in variable scale (for $X_i$ continuous) and number of levels (for $X_i$ discrete) introduces a
    bias. A uninformative variable with a large number of levels may be selected over a more informative variable with
    fewer levels;
  \item bagging introduces a bias, but sampling $(1- 1/e) \approx 0.632$ of the data without replacement seems to be an
    effective stragey for removing the problem.
  \end{enumerate}
\end{itemize}

\cite{Strobl.et.al.2007} introduce changes to the random forest algorithm to fix 1) and 2).  However the level effect will be
apparent in the CF algorithm as well.  However, there are many examples, particularly in genomics, where we have data
sets that are
\begin{itemize}
\item very wide;
\item all variables have the same number of levels (for example, measuments at different points along the genome);
\item the hypothesized functional relationship is of a form amenable to a decision tree.
\end{itemize}

@@

Compared to Spark ML, \cursedforest\ further parallelizes work down to individual features.  For example, for each node of
a tree, \cursedforest\ will distribute tasks that consist of single features (variants), for every individual.  Each of
these tasks will calculate the information gain for that specific feature.  Once these tasks have completed, the results
are reduced to return the feature which gives the greatest information gain.  This process is then repeated until
\cursedforest\ has calculated the entire decicion tree.

The current implementation of \cursedforest\ uses:
\begin{itemize}
\item an ``information gained'' criteria for splitting. Let $f_q$ be the fraction of items labeled with value $q$ where $q= 1,
  \ldots, Q$ at a node. The entropy of the node is $H = - \sum^{m}_{i=1} f_i \log^{}_2 f_i.$
The expected information gain is the change in entropy $H$ from a prior state $T$ to a state that takes some
information $a$ as given: 
\[ IG(T,a) = H(T) - H(T|a). \]
In this case, the information gain is IG = Entropy(parent node) - Weighted Sum of Entropy(Children)
\item variable importance is the sum of IG over all notes and all trees for that variable
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This section talks about the datasets
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Datasets}
%DENIS: Aidan can you please fill this in of where we got it from and how we process it for all the different tools to use
\subsubsection{1000 Genomes Project}

We obtained the 1000 Genomes Project data as VCF files from their FTP site.

%DENIS: Rob how was the synthetic data generated
\subsubsection{Synthetic data}

Each dataset consists of $n$ samples and $m$ variables where $n << m$, and values for each variable are ordinal factors with
three levels represented as numbers $\{0, 1, 2\}$ (which correspond to an additive effect encoding of genomics variation) randomly
generated from uniform distribution with equal probabilities.  The dichotomous response is generated from a two level
Bernoulli trail with $p$ (probability of success) calculated by applying inverse logit transformation to the continuous
linear response ($y$) of five variables:
$$y = a_{10}* v_{10} + a_{100}*v_{100} + a_{1000}*v_{1000} + a_{10000}*v_{10000} + a_{10000}*v_{100000}$$
where $a_{10} = 0.8, a_{100} = 0.9, a_{1000} = 1.1, a_{10000} = 1.0, a_{10000}=1.15$.

@@

\begin{itemize}
\item \ntree\ the number of trees.  The number of trees necessary for good performance grows with the number of predictors.
  \marginpar{can we do proximity?} \cite{Liaw.and.Weiner.2002} suggest that a high \ntree\ is necessary to get stable
  estimates of variable importance and proximity; however, even though the variable importance measures may vary from run to run,
  the ranking of the importances is quite stable.  In addition, we note that it is possible for a random forest model to have a very poor
  fit, and be useless for prediction, and still have a useful ranking of variable importance.
\item \mtry\ -- the number of variables considered at each split (if \mtry=$p$, we have a boosted decision
  tree model).  If one has a very large number of variables but expects only very few to be ``important'', using larger \mtry\ may give
  better performance.
\item For classification problems where the class frequencies are extremely unbalanced it may be necessary to change the prediction
  rule \marginpar{can we do this}
\item the size and complexity of the individual trees is controlled in {\sc randomForest} by setting \texttt{nodesize}, the
  minimum size of terminal nodes. It is controlled in Spark ML by setting \texttt{maxDepth}, the maximum depth of each
  tree in the forest. The Spark ML documentation somewhat unhelpfully sets \texttt{maxDepth=4} in their classification
  example\footnote{The documentation is available at \url{http://spark.apache.org/docs/latest/mllib-ensembles.html}.}.

  In the ATCG example the RandomForests code produced trees with depths of 30+. We note that random forests are a way of averaging
  multiple deep decision trees, trained on different parts of the same training set, with the goal of reducing the
  variance. \url{https://en.wikipedia.org/wiki/Random_forest} and refs therein.  
\marginpar{are we including this example?  Supplementary info?}
\end{itemize}

@@

\begin{figure}[tbhp]
\begin{tabular}{ll}
a)& b)\\
\includegraphics[totalheight=6cm]{./figs/var_oder.png} & 
\includegraphics[totalheight=6cm]{./figs/var_oob.png} \\
\end{tabular}
\captionof{figure}{{\bf Variable importance (Mean Decrease Gini) and OOB (out of bag) classification error estimate for
  random forests with 100 trees trained on the synthetic dataset of 2.5M features and 5K samples.}  {\bf a)} Variable
  importance of top 20 variabels normalized to the sum of the linear model coefficients. {\bf Others} represents
  aggregated importance of variables not present in the linear model.  {\bf b)} OOB (out of bag) classification error
  estimate as funtion of \mtry\ fraction.}
\label{figure:synth}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Section 2: Feature Selection on real data 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\cursedforest\ selects previously demonstrated cancer variants }
%Feature selection can be useful for determining variables the contribute to the label or for perhaps selecting the n most important
%variables for further analysis. 
As established in the pervious section, \cursedforest\ allows the user to retrieve a subsection of features, ranked by
their importance to identify the variables predictive of the label or selecting the $n$ most important variables for
further analysis.  We demonstrate this feature on The Cancer Genome Atlas (TCGA) BRCA dataset.  We build this model on
VCF files retrieved from the TCGA data portal.  For the label, we use the cell type (i.e. tumor or normal).

@@

\begin{table}[!ht]
\centering
\caption{
{\bf Time in seconds to build a random forest of 100
  trees with specified \mtry\ fraction using given number of CPU cores.}}
\begin{tabular}{|l|l||l|l|l|l|}
\hline
\multicolumn{2}{|l|}{\multirow{2}{*}{}}             & \multicolumn{4}{c|}{Number of CPU cores} \\
\cline{3-6}
\multicolumn{2}{|l|}{}                              & \bf{16} & \bf{32} & \bf{64} & \bf{128} \\
\hline
\multirow{4}{*}{\mtry\ (fraction)}        & \bf{0.10} & 708.3   & 340.0   & 186.0   & 134.5 \\
                                        & \bf{0.25} & 1134.9  & 535.0   & 270.8   & 179.6 \\
                                        & \bf{0.50} & 1724.1  & 856.6   & 428.7   & 243.8 \\
                                        & \bf{0.80} & 2313.5  & 1125.7  & 569.1   & 364.6 \\
\hline
\end{tabular}
\begin{flushleft} 
\end{flushleft}
\label{table10}
\end{table}

\begin{figure}[tbhp]
\begin{tabular}{ll}
a)& b)\\
\includegraphics[totalheight=6cm]{./figs/mtry_cpu.png} & 
\includegraphics[totalheight=6cm]{./figs/cpu_mtry_trees_per_hour.png} \\
\end{tabular}
\captionof{figure}{{\bf Scalablity of the wide random forest on the synthetic dataset of 2.5M features and 5K samples.} 
  {\bf a)} Time in seconds to build 100 trees for different \mtry\ fractions. 
  {\bf b)} Number of trees build per hour when using a growing number of CPU cores.}
\label{figure:synth}
\end{figure}

@@

\begin{table}[!ht]
\centering
\caption{
{\bf Time in seconds to build a random forest of 100
  trees on 128 CPU cores with varying numbers of variables and samples.}}
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
\multicolumn{2}{|l|}{\multirow{2}{*}{}}           & \multicolumn{5}{c|}{Number of variables} \\
\cline{3-7}
\multicolumn{2}{|l|}{}                               & \bf{150,000} & \bf{500,000} & \bf{2,500,000}  & \bf{10,000,000} & \bf{50,000,000} \\
\hline                                                            
\multirow{4}{*}{Number of samples}      & \bf{1000}  & 21.6  & 26.7  & 59.8  & 133.5  & 668.8 \\
                                        & \bf{5000}  & 45.7  & 71.9  & 186.2 & 729.3  & 2887.4 \\
                                        & \bf{10000} & 70.2  & 120.2 & 397.3 & 2237.8 & 5747.8 \\
\hline
\end{tabular}
\begin{flushleft} 
\end{flushleft}
\label{table10}
\end{table}

\begin{figure}[tbhp]
\begin{tabular}{ll}
a)& b)\\
\includegraphics[totalheight=6cm]{./figs/nvars_nsamples.png} & 
\includegraphics[totalheight=6cm]{./figs/nsamples_nvars.png} \\
\end{tabular}
\captionof{figure}{{\bf Scalablity of the Wide Random Forest on synthetic datasets with varying number of samples and variables.} 
  {\bf a)} Number of trees build per hour when growing number of variables. 
  {\bf b)} Number of trees build per hour when growing number of samples.}
\label{figure:synth}
\end{figure}

\begin{table}[!ht]
%\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
\caption{
{\bf Sensitivity analysis on parameter choice.}}
\begin{tabular}{|l|l|l|l|l|l|l|l|}
\hline
%\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \hline
\bf{ \mtry\ }  & \bf{Mtree} & \bf{depth} & \bf{Accuracy} & \bf{Runtime} & \bf{Memory} \\
\hline
&&&&&\\ \hline
\end{tabular}
\begin{flushleft} 
  Table notes here.
\end{flushleft}
\label{table2}
%\end{adjustwidth}
\end{table}

>>>>

\end{flushleft}
\tableofcontents
\clearpage
\section{Notes}
\begin{itemize}
\item I think the structure of the first section should be
  \begin{itemize}
  \item convergence results from Biau 2012  (done)
  \item consider  \cite{Segal.2004} examples
  \item simulations reproduced from Genuer et al 2010 (no -- we need classification examples)
  \item check Piotr's simulation  (done)
  \item run spark ML or CursedForest on the  simulation data 
  \item we extend the simulations to many more variables 
  \item I dont think I can get R to handle chromosme 1 etc. May have to run CF on chromosme 22, see table
    \ref{table:smaller_data} -- On it, Aidan
  \item variable  selection may succeed where accurate prediction does not
  \item discussion of the effect of the \ntree\ and \mtry\ parameters for variable selection with many noise variables
  \item   perhaps we argue that in the case where the relationship of $X$ and $y$ takes a particular form (step, ramp) we can recover
    the significant variables in the presence of a very large number of noise variables
  \item in the case of genomic data, our functions of interest are more likely to be stop or ramp functions rather than harmonic
    functions. So we have some hope.
  \end{itemize}
\item check out 
  \begin{itemize}
  \item D\'iaz-Uriarte, R., \& Alvarez de Andr\'es, S. (2006). (done)
  \item \cite{Goldstein.et.al.2011}
notes that the 0, 1, 2 encoding of alleles is an additive model i.e. each additional minor allele increases the effect 
\begin{verbatim}
Different genetic effects
Type	Mechanism	Partition
Additive	Each additional minor allele increases variation	0, 1, 2
Dominant	Presence of at least 1 minor allele increases variation	0, 1/2
Recessive	Two minor alleles needed for variation	0/1, 2
Heterosis	Heterozygote leads to variation	0/2, 1
\end{verbatim}

@@

\marginpar{decison trees capture interaction -- there is no main effect term. Random forests inherit this property. I dont think it
  has anything to do with boostrapping, the point of which is to ensure that successive trees are not correlated.}
Random forests~\cite{Breiman2001} are able to capture interactions between features while still allowing to
prioritise SNPs for their contribution to disease.  The joint interaction is achieved by applying the technique of
bootstrap aggregation or bagging to decision trees, while a measure of variable importance can be provided, for example,
by summing the number of times a variable is used to split a node, over all nodes and all trees.  Random forests are well suited for
processing ``wide'' genomic data for two reasons.  Firstly, while other machine learning applications have the
propensity to overfit datasets with more features $p$ than samples $n$ (a consequence of the ``curse of
dimensionality''~\cite{Bauer2014, bellman1961adaptive}), decision trees are resistant to overfitting. 
Secondly, random forests are also very easy to parallelise. As the forest is a sum of decision trees, it it possible to grow
separate trees on different processors and combine the trees.

The most popular random forest software packages are R-based, although current implementations do not scale well with
increasing feature size~\cite{Wright.and.Ziegle.2016}.  This is despite R now supporting long vectors (length greater
than $2^{31}$), and the {\sc randomForest} implementation in R lending itself to parallelization to grow multiple trees
simultaneously~\cite{Liaw.and.Weiner.2002} and them combine them.  More suitable implementations for ``wide'' genomic data have been
developed in other programming languages.  {\sc CloudForest}~\cite{Bressler2015} written in Go, achieves fast running
times by effective use of the CPU cache, optimising for different classes of features and efficiently multi-threading.
Similarly, {\sc Ranger}~\cite{Wright.and.Ziegle.2016} written in C++ with an R front-end is specifically optimised for
dealing with large data sets.

@@

Since the original implementation of random forests, there have been a number of developments leading to a better understanding of
the algorithm. 
\begin{itemize}
\item random forests are resistant to overfitting. However it is not true that they will not overfit,
  ~\cite{Segal.2004} gives examples of overfitting in the case where:
  \begin{itemize}
  \item a single tree is the most appropriate model; and
  \item the variables are highly correlated so that the bootstrap resampling does not give rise to uncorrelated trees.
  \end{itemize}
However, keeping these difficult cases in mind, random forests have a strong claim to being resistant to overfitting.

\item \cite{Strobl.et.al.2007} demonstrates that the random forest algorithm is subject to a bias in variable selection via two mechanisms:
  \begin{enumerate}
  \item differences in variable scale (for $X_i$ continuous) and number of levels (for $X_i$ discrete) introduces a
    bias. A uninformative variable with a large number of levels may be selected over a more informative variable with
    fewer levels;
  \item bagging introduces a bias, but sampling $(1- 1/e) \approx 0.632$ of the data without replacement seems to be an
    effective strategy for removing the problem.
  \end{enumerate}
\end{itemize}

\cite{Strobl.et.al.2007} introduce changes to the random forest algorithm to fix 1) and 2).  However the level effect will be
apparent in the CF algorithm as well.  However, there are many examples, particularly in genomics, where we have data
sets that are
\begin{itemize}
\item very wide;
\item all variables have the same number of levels (for example, measurements at different points along the genome);
\item the hypothesized functional relationship is of a form amenable to a decision tree.
\end{itemize}

@@

Compared to Spark ML, \cursedforest\ further parallelizes work down to individual features.  For example, for each node of
a tree, \cursedforest\ will distribute tasks that consist of single features (variants), for every individual.  Each of
these tasks will calculate the information gain for that specific feature.  Once these tasks have completed, the results
are reduced to return the feature which gives the greatest information gain.  This process is then repeated until
\cursedforest\ has created the entire decision tree.

The current implementation of \cursedforest\ uses:
\begin{itemize}
\item an ``information gained'' criteria for splitting. Let $f_q$ be the fraction of items labelled with value $q$ where $q = 1,
  \ldots, Q$ at a node. The entropy of the node is $H = - \sum^{m}_{i=1} f_i \log^{}_2 f_i.$
The expected information gain is the change in entropy $H$ from a prior state $T$ to a state that takes some
information $a$ as given: 
\[ IG(T,a) = H(T) - H(T|a). \]
In this case, the information gain is IG = Entropy(parent node) - Weighted Sum of Entropy(Children)
\item variable importance is the sum of IG over all notes and all trees for that variable
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This section talks about the datasets
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Datasets}
\subsubsection{1000 Genomes Project}
We obtained the 1000 Genomes Project data as VCF files from their FTP site.  Each VCF file contains the variants for every individual accross one chromosome.  
We use the phase 3 dataset which contains 2,504 individuals with approximately 81 million variant positions (i.e. a matrix of $2,504 \times 81$ million).  
For compatibility with the majority of the algorithms we demonstrate, we convert the original variant representations to double representations.  
For example, we store the absence of a variant (0\textbar0) as 0.0, a heterozygous variant (1\textbar0 or 0\textbar1)as 1.0. and a homozygous variant (1\textbar1) as 2.0.

%DENIS: Rob how was the synthetic data generated
\subsubsection{Synthetic data} 
Each dataset consists of $n$ samples and $m$ variables where $n << m$, and values for each variable are ordinal factors with
three levels represented as numbers $\{0, 1, 2\}$ (which correspond to an additive effect encoding of genomics variation) randomly
generated from uniform distribution with equal probabilities.  The dichotomous response is generated from a two level
Bernoulli trail with $p$ (probability of success) calculated by applying inverse logit transformation to the continuous
linear response ($y$) of five variables:
$$y = a_{10}* v_{10} + a_{100}*v_{100} + a_{1000}*v_{1000} + a_{10000}*v_{10000} + a_{10000}*v_{100000}$$
where $a_{10} = 0.8, a_{100} = 0.9, a_{1000} = 1.1, a_{10000} = 1.0, a_{10000}=1.15$.

@@

\begin{itemize}
\item \ntree\ the number of trees.  The number of trees necessary for good performance grows with the number of predictors.
  \marginpar{can we do proximity?} \cite{Liaw.and.Weiner.2002} suggest that a high \ntree\ is necessary to get stable
  estimates of variable importance and proximity; however, even though the variable importance measures may vary from run to run,
  the ranking of the importances is quite stable.  In addition, we note that it is possible for a random forest model to have a very poor
  fit, and be useless for prediction, and still have a useful ranking of variable importance.
\item \mtry\ -- the number of variables considered at each split (if \mtry=$p$, we have a boosted decision
  tree model).  If one has a very large number of variables but expects only very few to be ``important'', using larger \mtry\ may give
  better performance.
\item For classification problems where the class frequencies are extremely unbalanced it may be necessary to change the prediction
  rule \marginpar{can we do this}
\item the size and complexity of the individual trees is controlled in {\sc randomForest} by setting \texttt{nodesize}, the
  minimum size of terminal nodes. It is controlled in Spark ML by setting \texttt{maxDepth}, the maximum depth of each
  tree in the forest. The Spark ML documentation sets \texttt{maxDepth} to 4 in their classification
  example\footnote{The documentation is available at \url{http://spark.apache.org/docs/latest/mllib-ensembles.html}.}.

  In the TCGA example the RandomForests code produced trees with depths of 30+. We note that random forests are a way of averaging
  multiple deep decision trees, trained on different parts of the same training set, with the goal of reducing the
  variance. \url{https://en.wikipedia.org/wiki/Random_forest} and refs therein.  
\marginpar{are we including this example?  Supplementary info?}
\end{itemize}

@@

\begin{figure}[tbph]
\begin{adjustwidth}{-1.00in}{0in}
\begin{tabular}{ll}
a)& b)\\
\includegraphics[totalheight=6cm]{./figs/var_oder.png} & 
\includegraphics[totalheight=6cm]{./figs/var_oob.png} \\
\end{tabular}
\caption{{\bf Variable importance (Mean Decrease Gini) and OOB (out of bag) classification error estimate for
  random forests with 100 trees trained on the synthetic dataset of 2.5M features and 5K samples.}  {\bf a)} Variable
  importance of top 20 variabels normalized to the sum of the linear model coefficients. {\bf Others} represents
  aggregated importance of variables not present in the linear model.  {\bf b)} OOB (out of bag) classification error
  estimate as funtion of \mtry\ fraction.}
\label{figure:synth}
\end{adjustwidth}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Section 2: Feature Selection on real data 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\cursedforest\ selects previously demonstrated cancer variants}
%Feature selection can be useful for determining variables the contribute to the label or for perhaps selecting the n most important
%variables for further analysis. 
As established in the pervious section, \cursedforest\ allows the user to retrieve a subsection of features, ranked by
their importance to identify the variables predictive of the label or selecting the $n$ most important variables for
further analysis.  We demonstrate this feature on The Cancer Genome Atlas (TCGA) BRCA dataset.  We build this model on
VCF files retrieved from the TCGA data portal.  For the label, we use the cell type (i.e. tumor or normal).

@@

\begin{table}[!ht]
\centering
\caption{
{\bf Time in seconds to build a random forest of 100
  trees with specified \mtry\ fraction using given number of CPU cores.}}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\multicolumn{2}{|l|}{\multirow{2}{*}{}}             & \multicolumn{4}{c|}{Number of CPU cores} \\
\cline{3-6}
\multicolumn{2}{|l|}{}                              & \bf{16} & \bf{32} & \bf{64} & \bf{128} \\
\hline
\multirow{4}{*}{\mtry\ (fraction)}        & \bf{0.10} & 708.3   & 340.0   & 186.0   & 134.5 \\
                                        & \bf{0.25} & 1134.9  & 535.0   & 270.8   & 179.6 \\
                                        & \bf{0.50} & 1724.1  & 856.6   & 428.7   & 243.8 \\
                                        & \bf{0.80} & 2313.5  & 1125.7  & 569.1   & 364.6 \\
\hline
\end{tabular}
\begin{flushleft} 
\end{flushleft}
\label{table10}
\end{table}

\begin{figure}[tbhp]
\begin{adjustwidth}{-1.00in}{0in}
\begin{tabular}{ll}
a)& b)\\
\includegraphics[totalheight=6cm]{./figs/mtry_cpu.png} & 
\includegraphics[totalheight=6cm]{./figs/cpu_mtry_trees_per_hour.png} \\
\end{tabular}
\caption{{\bf Scalablity of the wide random forest on the synthetic dataset of 2.5M features and 5K samples.} 
  {\bf a)} Time in seconds to build 100 trees for different \mtry\ fractions. 
  {\bf b)} Number of trees build per hour when using a growing number of CPU cores.}
\label{figure:synth}
\end{adjustwidth}
\end{figure}

@@

\begin{table}[!ht]
\begin{adjustwidth}{-1.00in}{0in}
\centering
\caption{
{\bf Time in seconds to build a random forest of 100
  trees on 128 CPU cores with varying numbers of variables and samples.}}
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
\multicolumn{2}{|l|}{\multirow{2}{*}{}}           & \multicolumn{5}{c|}{Number of variables} \\
\cline{3-7}
\multicolumn{2}{|l|}{}                               & \bf{150,000} & \bf{500,000} & \bf{2,500,000}  & \bf{10,000,000} & \bf{50,000,000} \\
\hline                                                            
\multirow{4}{*}{Number of samples}      & \bf{1000}  & 21.6  & 26.7  & 59.8  & 133.5  & 668.8 \\
                                        & \bf{5000}  & 45.7  & 71.9  & 186.2 & 729.3  & 2887.4 \\
                                        & \bf{10000} & 70.2  & 120.2 & 397.3 & 2237.8 & 5747.8 \\
\hline
\end{tabular}
\begin{flushleft} 
\end{flushleft}
\label{table10}
\end{adjustwidth}
\end{table}

\begin{figure}[tbhp]
\begin{adjustwidth}{-1.00in}{0in}
\begin{tabular}{ll}
a)& b)\\
\includegraphics[totalheight=6cm]{./figs/nvars_nsamples.png} & 
\includegraphics[totalheight=6cm]{./figs/nsamples_nvars.png} \\
\end{tabular}
\caption{{\bf Scalablity of the Wide Random Forest on synthetic datasets with varying number of samples and variables.} 
  {\bf a)} Number of trees build per hour when growing number of variables. 
  {\bf b)} Number of trees build per hour when growing number of samples.}
\label{figure:synth}
\end{adjustwidth}
\end{figure}

\begin{table}[!ht]
\caption{
{\bf Sensitivity analysis on parameter choice.}}
\begin{tabular}{|l|l|l|l|l|l|l|l|}
\hline
%\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \hline
\bf{ \mtry\ }  & \bf{Mtree} & \bf{depth} & \bf{Accuracy} & \bf{Runtime} & \bf{Memory} \\
\hline
&&&&&\\ \hline
\end{tabular}
\begin{flushleft} 
  Table notes here.
\end{flushleft}
\label{table2}
\end{table}
